---
title: "Human Activity Recognisation"
author: "Wong Man Hin"
date: "2015-12-25"
output: html_document
---
##Introduction
In this article, we attempted to build a classification model to recognise human activity through motion information gathered by devices, using random forest technique.

##Data preparation
Packages used in the analysis is imported.
```{r cache=TRUE}
library(dplyr)
library(caret)
```

Data is loaded into R for further processing. Only readings of Eular Angles and readings from accelerometer, gyroscope and magnetometer of the 3 devices are used in the model training.

```{r cache =TRUE}
train <- read.csv("pml-training.csv", stringsAsFactors = FALSE)
train_information <- select(train, matches("^roll|^pitch|^yaw|^gyros|^accel|^magnet|classe"))
```

Training dataset will be divided into 2 parts, 70% of the data will be training set, while the remaining 30% will be the testing set.

```{r cache=TRUE}
trainIndex <- createDataPartition(train_information$classe, p = 0.70,list=FALSE)
training <- train_information[trainIndex,]
validation <- train_information[-trainIndex,]
```

##Model training
Caret package is in the model training process. Random forest method with number of tree = 100 is used as the training method. 3-fold cross validation is performed to reduce the probability of overfitting.
```{r cache=TRUE}
rf_model <- train(classe~., data = training, method = "rf", trcontrol = trainControl(method = "cv", number = 3),allowParallel = TRUE, ntree = 100)
print(rf_model$finalModel)
```

From the final Model, we get a accuracy rate of `r (1-rf_model$finalModel$err.rate[nrow(rf_model$finalModel$err.rate),1])*100`%

##Model testing
After building a random forest model for the classification, the model is tested by the testing set.

Model is used to perform the prediction of the testing set.
```{r cache=TRUE}
validation_prediction <- predict(rf_model, newdata = validation)
confusionMatrix(validation_prediction,validation$classe)
```


Model is validated and comfortable to predict other cases. Out of sample error is expected to be 1%